# Analysis of acitivity monitoring data

========================================================

### Loading and preprocessing the data

1.Load the data.
```{r}
data <- read.csv("./activity.csv")
```

### Mean total number of steps taken per day

1.Calculate total number of steps taken each day.
```{r}
s <- split(data, data$date)
totalSteps <- lapply(s, function(x) sum(x[,1], na.rm = TRUE))
totalSteps <- data.frame(names(totalSteps), unlist(totalSteps))
rownames(totalSteps) <- 1:nrow(totalSteps)
names(totalSteps) <- c("date","steps")
```

2.Make a histogram of the total number of steps taken each day.
```{r}
hist(totalSteps$steps)
```

3.The mean and median total number of steps taken per day.
```{r}
mean(totalSteps$steps)
median(totalSteps$steps)
```

### What is the average daily acitivity pattern?

1.Calculate the average number of steps at the same time of every day.
```{r}
s2 <- split(data, data$interval)
averagedSteps <- lapply(s2, function(x) mean(x[,1], na.rm = TRUE))
averagedSteps <- data.frame(names(averagedSteps), unlist(averagedSteps))
rownames(averagedSteps) <- 1:nrow(averagedSteps)
names(averagedSteps) <- c("interval","steps")
averagedSteps$interval <- as.numeric(levels(averagedSteps$interval))[averagedSteps$interval]
```

2.Make a time series plot.
```{r}
plot(averagedSteps$interval, averagedSteps$steps, type = "l", col = "red", main = "5-minute interval and the average number of steps taken", xlab = "Interval", ylab = "Number of steps")
```

3.The 5-minute interval that contains the maximum number of steps.
```{r}
averagedSteps[averagedSteps$steps == max(averagedSteps$steps),][,1]
```

### Imputing missing values

1.Calculate the total number of missing values in the dataset.
```{r}
nrow(data[is.na(data[,1]),])
```

2.Create a new dataset and a subset that contains missing values. Fill in the missing values using the mean of the 5-minute interval.
```{r}
data2 <- data
subset <- data2[is.na(data2[,1]),]
 for(i in rownames(subset))
         { 
         data2[as.numeric(i),1] <-  averagedSteps[averagedSteps$interval== data2[as.numeric(i),3],2] 
         }
```

3.Make a new histogram of the total number of steps taken each day.
```{r}
s3 <- split(data2, data2$date)
totalSteps2 <- lapply(s, function(x) sum(x[,1], na.rm = TRUE))
totalSteps2 <- data.frame(names(totalSteps2), unlist(totalSteps2))
rownames(totalSteps2) <- 1:nrow(totalSteps2)
names(totalSteps2) <- c("date","steps")
hist(totalSteps2$steps)
```

4.Calculate the mean and median total number of steps taken per day after imputing missing values.
```{r}
mean(totalSteps2$steps)
median(totalSteps2$steps)
```

These two values don't differ from the first part of the assignment. The impact of imputing missing data can be ignored.

### Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels "weekday" and "weekend".
```{r}
data2$weekday<-factor(weekdays(as.Date(data2$date)), levels = c("Monday","Tuesay","Wednesday","Thursday","Friday","Saturday","Sunday","weekday","weekend"))
data2$weekday[data2$weekday=="Monday"] <- factor("weekday")
data2$weekday[data2$weekday=="Tuesday"] <- factor("weekday")
data2$weekday[data2$weekday=="Wednesday"] <- factor("weekday")
data2$weekday[data2$weekday=="Thursday"] <- factor("weekday")
data2$weekday[data2$weekday=="Friday"] <- factor("weekday")
data2$weekday[data2$weekday=="Saturday"] <- factor("weekend")
data2$weekday[data2$weekday=="Sunday"] <- factor("weekend")
```

2.Subset and split the dataset based on the factor.
```{r}
subset2 <- data2[data2$weekday=="weekday",]
s4 <- split(subset2, subset2$interval)
subset3 <- data2[data2$weekday=="weekend",]
s5 <- split(subset3, subset3$interval)
```

3.Calculate the average number of steps taken for each subset.And then combine these two subsets together. 
Make a panel plot containing a time series plot.
```{r}
avewd <- lapply(s4, function(x) mean(x[,1]))
avewd <- data.frame(names(avewd),unlist(avewd))
rownames(avewd) <- 1:nrow(avewd)
names(avewd) <- c("interval","steps")
avewd$interval <- as.numeric(levels(avewd$interval))[avewd$interval]
avewd$weekday <- factor("weekday")
```

```{r}
avewe <- lapply(s5, function(x) mean(x[,1]))
avewe <- data.frame(names(avewe),unlist(avewe))
rownames(avewe) <- 1:nrow(avewe)
names(avewe) <- c("interval","steps")
avewe$interval <- as.numeric(levels(avewe$interval))[avewe$interval]
avewe$weekday <- factor("weekend")
ave <- rbind(avewd,avewe)
library(lattice)
xyplot(ave$steps ~ ave$interval|ave$weekday, layout = c(1,2), type = "l", xlab = "Interval", ylab = "Number of steps")
```
